---
title: "rnorm for n00b"
Author: Lucien De Brot
output: 
  html_document: 
    css: n00b.css
---

## RNORM - Random Normal Distribution

**rnorm is useful for creating a set of normally-distributed observations.**

**rnorm's arguments**:

* n = number of observations to be created (=size of the distribution)

* mean = mean of the distribution

* sd = standard deviation of the distribution

**rnorm**(**x**=*number of values*, **mean**=*mean*, **sd**=*standard deviation*)

Let us generate such a distribution with 10000 observations, centered around the mean 174 and with a standard deviation of 10.

```{r generate_test_pop, echo=TRUE}
test_pop <- rnorm(n=10000,mean=174,sd=10)
plot(test_pop)
```

* On the x-axis, you find the number of each observation (value number one, value number two, etc.).

* On the y-axis, you find the actual value of the normally-distributed 

In order to see that the newly created set of observations is actually normally-distributed, we need to plot a histogram.

```{r plot_test_pop, echo=TRUE}
hist(test_pop,breaks=100)
```

This shows the nice "bell form" of a normal distribution.

```{r qqnorm_test_pop, echo=TRUE}
qqnorm(test_pop)
qqline(test_pop,col="red")
```

And the **normal probability plot** above confirms the normal distribution graphically.
