---
title: "Confidence intervals"
author: "Lucien De Brot"
date: "25 juillet 2016"
output: 
  html_document: 
    css: C:/Users/Lucien/OneDrive/HARDDRIVE/LDS/NotAnEngineer/n00b/n00b.css
---

#Before calculating the confidence interval

Criteria to find the right formulae:

* Distribution : normal, poisson, t,...?

* Independent or paired data?

* One or two means?

* Equal or different variance?

* one- or two-sided? (draw a picture)

#Normal Confidence intervals

```{r norm_conf_int}
#HYPOTHESES
#H_0: mean - mean_old = 0
#H_a: mean - mean_old > 0
#two-sided
#test statistic : Z score
#n > 30, randomized, 

#GIVEN DATA
n <- 64 ### patients
mean <- 137.5 ### minutes
sd <- 39 ### minutes
mean_old <- 127 ### minutes
alpha <- .99
sides <- 2 ### two-sided
prob <- (1+alpha)/sides

#CALCULATED DATA
SE <- sd/sqrt(n)
z_score <- (mean-mean_old)/SE
confidence_interval <- mean + c(-1,1) * qnorm(p=prob) * SE
p_value <- round((1-pnorm(z_score))*sides,4)
```

#T CONFIDENCE INTERVAL

```{r t_conf_int}
#In a population of interest, a sample of 9 men yielded a sample average brain volume of 1,100cc and a standard deviation of 30cc. What is a 95% Student's T confidence interval for the mean brain volume in this new population?

#GIVEN DATA
n <- 9
mean <- 1100
sd <- 30
alpha <- .05
sides <- 2 ### two-sided
prob <- (1+(1-alpha))/sides

#CALCULATED DATA
SE <- sd/sqrt(n)
df <- n-1
confidence_interval <- mean + c(-1,1) * qt(p=prob,df) * SE
```

#PAIRED GROUP T INTERVALS

``` {r iprd_grp_t_int}
#GIVEN DATA
n_diff <- 73 ### Number of "data pairs"
mean_diff <- 12.76 ### Mean of all differences
sd_diff <- 14.26 ### standard deviation of the sample of differences

#CALCULATED DATA
SE_diff <- sd_diff/sqrt(n_diff)
t_score <- (mean_diff - 0)/SE_diff
```

#INDEPENDENT GROUP T INTERVALS (equal variance / pooled standard deviation)

``` {r ind_grp_t_int_eq_var}
#GIVEN DATA
n <- 20
n_t <- 10
n_c <- 10
mean_t <- 3
mean_c <- 5
sd_t <- sqrt(.6)
sd_c <- sqrt(.68)

#CALCULATED DATA
var <- ((n_t-1)*sd_t^2+(n_c-1)*sd_c^2)/df ### For pooled variance only 
sd <- sqrt(var)
SE <- sd*sqrt(1/n_t+1/n_c) ### For pooled standard deviation only
df <- n_t + n_c - 2
difference <- mean_t - mean_c
confidence_interval <- difference + c(-1,1) * qt(.975,df) * SE
```

#INDEPENDENT GROUP T INTERVALS (unequal variance)

``` {r ind_grp_t_int_uneq_var}
#GIVEN DATA
n <- 20
n_t <- 10
n_c <- 10
mean_t <- 3
mean_c <- 5
var_t <- .6
var_c <- .68

#CALCULATED DATA
sd_t <- sqrt(var_t)
sd_c <- sqrt(var_c)
SE <- sqrt(var_t/n_t+var_c/n_t) ### For unequal variances (vs pooled ones)
df <- (sd_t^2/n_t+sd_c^2/n_c)^2/(((sd_t^2/n_t)^2/(n_t-1))+((sd_c^2/n_c)^2/(n_c-1))) ### For unequal variances (vs pooled ones)
difference <- mean_t - mean_c
confidence_interval <- difference + c(-1,1) * qt(.975,df) * SE

#POWER
effect <- -3 ## The effect of interest is a -3 difference between the treatment and the control group
upun_confint <- difference + (effect/abs(effect)) * qt(.975,df) * SE
power <- pnorm((upun_confint - effect)/SE)

```

#POISSON INTERVAL

``` {r poisson}
#GIVEN DATA
lambda <- 25
alpha <- .05
sides <- 2
t <- 3

#CALCULATED DATA
SE <- sqrt(lambda/t)
confidence_interval <- lambda + c(-1, 1) * qnorm(1-(alpha/sides)) * SE
```

#Clear everything

```{r clear_everything}
n <- NULL
mean <- NULL
sd <- NULL
alpha <- NULL
sides <- NULL
prob <- NULL
SE <- NULL
df <- NULL
confidence_interval <- NULL
```
