
---
title: "Distributions and probability"
author: "Lucien De Brot"
date: "21 juillet 2016"
output: 
  html_document: 
    css: C:/Users/Lucien/OneDrive/HARDDRIVE/LDS/NotAnEngineer/n00b/n00b.css
---

#Normal distribution
Randomized sample of independent and identically distributed variables.

#Geometric distribution
The first success in *n* Bernoulli trials

#Binomial distribution
The *k*th success in the *n*th trial

Check whether the distribution is binomial

```{r binom_check, echo=TRUE}
binom_check <- function () {
    print("1) The trials are independent")
    print("2) The number of trials *n* is fixed")
    print("3) Each trial outcome can be classified as a *success* or *failure*")
    print("4) The probability of success *p* is the same for each trial")
}
```

Obtain the probability of *k* successes in *n* trials at probability *p*:

```{r binom_probability, echo=TRUE}
binom_probability <- function (n=0,k=0,p=0) {
    if (sum(c(n,k,p))==0) {
        print("You need to specify n trials, k successes and p probability")
        print("FORMULA : choose(n,k)*p^k*(1-p)^(n-k)")
    } else {
        choose(n,k)*p^k*(1-p)^(n-k)
    }
}
```

Obtain the mean of a binomial distribution:

```{r binom_mean, echo=TRUE}
binom_mean <- function (n=0,p=0) {
    if (sum(c(n,p))==0) {
        print("You need to specify n trials and p probability.")
        print("FORMULA : Mean = n * p*")
    } else {
        print(n*p)
    }
}
```

Obtain the standard deviation of a binomial distribution:

```{r binom_sd, echo=TRUE}
binom_sd <- function (n=0,p=0) {
    if (sum(c(n,p))==0) {
        print("You need to specify n trials and p probability.")
        print("FORMULA : sigma = sqrt(n*p*(1-p))")
    } else {
        print(sqrt(n*p*(1-p)))
    }
}
```

#Negative binomial distribution

Check whether the distribution is negative binomial

```{r nbinom_check, echo=TRUE}
nbinom_check <- function () {
    print("1) The trials are independent")
    print("2) Each trial outcome can be classified as a *success* or *failure*")
    print("3) The probability of success *p* is the same for each trial")
    print("4) The last trial is a success")
}
```

Obtain the probability of the *k*th success in *n* trials at probability *p*:

```{r nbinom_probability, echo=TRUE}
nbinom_probability <- function (n=0,k=0,p=0) {
    if (sum(c(n,k,p))==0) {
        print("You need to specify k successes at the nth trial, at p probability")
        print("FORMULA : choose(n-1,k-1)*p^k*(1-p)^(n-k)")
    } else {
        print(choose(n-1,k-1)*p^k*(1-p)^(n-k))
    }
}
```


#Poisson distribution

The Poisson distribution is useful for estimating the number of events in a large population over a unit of time.

Check whether it is a Poisson distribution:

```{r pois_check, echo=TRUE}
pois_check <- function () {
    print("1) We are looking for the number of events (=successes)")
    print("2) The population is large")
    print("3) Events occur independently from each other")
}
```

Obtain Poisson probability:

```{r pois_probability, echo=TRUE}
pois_probability <- function (lambda=0,k=0) {
    if (sum(c(lambda,k))==0) {
        print("You need to specify lambda and k successes")
        print("Note the difference between pois_probability and ppois! pois_probability provides probability for EXACTLY k successes, whereas ppois provides probability of k or less successes (=chunk of the distribution)")
        print("FORMULA : ((lambda^k)*exp(1)^(-1*lambda))/factorial(k)")
    } else {
        print(((lambda^k)*exp(1)^(-1*lambda))/factorial(k))
    }
}
```



Normal model can be used for sampling distributions if
* sample size > 30
* independent probabilities
* randomized


